---
title: "Question_2__drivers_of_heterogeneity"
author: "Kendra Walters"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
.main-container { max-width: 1000px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

---

# Purpose
To test whether plant composition is the main driver of heterogeneity (or, other option, spatial distance which may stand in for other factor(s)) AND, if it is, then to test at what radius do we see the strongest impact of plant composition. 

```{r setup, message=FALSE}
library(knitr)
library(tidyverse)
library(data.table)
library(kableExtra)
require(ggplot2)
require(gridExtra)
library(car)
require(vegan)
require(magrittr)
library(indicspecies)
require(pairwiseAdonis)
require(ncf)
require(conflicted)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")


knitr::opts_knit$set(root.dir = "../")

knitr::opts_chunk$set(comment="", cache=T, warning = F, message = F, fig.path = "05_output_figures/")

```

## Plants v distance
It's like Plants versus Zombies but much less interactive. Basically, we are doing a partial mantel here to see if geographic distance and/or plant composition is driving community composition in the bacterial and fungal dispersal communities found on the glass slides. For this analysis, we are arbitrarily using the plant composition within 1 meter of each site. We could have chosen a different radius here, but this one seemed OK.

```{r distance_v_plants__bacteria}
print("For BACTERIA") # just so we know what the output is for in the HTML file


## MICROBIAL MATRIX
# Clean up the Bray-Curtis matrix for bacterial communities
load("03_Processed_data/16S_glass_slides.rda")
M.microbial <- q1.bray.dist.df %>% 
  select(starts_with("LO")) %>% 
  rename(LOT1A3 = LOT1A3_dup_new_protocol) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  column_to_rownames("SampleID") %>% as.matrix


## DISTANCE MATRIX
# Load site coordinates and use to make dataframe of coordinates for each sample
site.coords <- as.data.frame(fread("03_Processed_data/site_locs_coordinates.tsv")) %>% 
  rename(Site = V1)

M.distance <- data.frame("Sample" = row.names(M.microbial)) %>% # start with same names in same order
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>%  # get site ID
  left_join(site.coords) %>% # join with our coordinates
  column_to_rownames("Sample") %>% 
  select(x,y) %>% dist(., method = "euclidean") %>% # create Euclidean distance matrix
  as.matrix


## PLANT MATRIX
## reading in plant composition bc and make df of plant composition of each sample

plants <- fread("03_Processed_data/map/grass_shrub_1_meter_composition_df.tsv", data.table = FALSE) %>% 
  rename(Site = V1)

M.plants <- data.frame("Sample" = row.names(M.microbial)) %>% 
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>% 
  left_join(plants) %>% 
  column_to_rownames("Sample") %>% 
  select(!(Site)) %>% 
  vegdist() %>% as.matrix



## Partial mantel time!!
partial.mantel.test(M.microbial, M.distance, M.plants, resamp = 1000, method = "pearson", quiet = TRUE)

```



```{r distance_v_plants__fungi}
print("For FUNGI") # just so we know what the output is for in the HTML file

## MICROBIAL MATRIX
# Clean up the Bray-Curtis matrix for bacterial communities
load("03_Processed_data/ITS_glass_slides.rda")
M.microbial <- q1.ITS.bray.dist.df %>% 
  select(starts_with("LO")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  column_to_rownames("SampleID") %>% as.matrix


## DISTANCE MATRIX
# Load site coordinates and use to make dataframe of coordinates for each sample
site.coords <- as.data.frame(fread("03_Processed_data/site_locs_coordinates.tsv")) %>% 
  rename(Site = V1)

M.distance <- data.frame("Sample" = row.names(M.microbial)) %>% # start with same names in same order
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>%  # get site ID
  left_join(site.coords) %>% # join with our coordinates
  column_to_rownames("Sample") %>% 
  select(x,y) %>% dist(., method = "euclidean") %>% # create Euclidean distance matrix
  as.matrix


## PLANT MATRIX
## reading in plant composition bc and make df of plant composition of each sample

plants <- fread("03_Processed_data/map/grass_shrub_1_meter_composition_df.tsv", data.table = FALSE) %>% 
  rename(Site = V1)

M.plants <- data.frame("Sample" = row.names(M.microbial)) %>% 
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>% 
  left_join(plants) %>% 
  column_to_rownames("Sample") %>% 
  select(!(Site)) %>% 
  vegdist() %>% as.matrix



## Partial mantel time!!
partial.mantel.test(M.microbial, M.distance, M.plants, resamp = 1000, method = "pearson", quiet = TRUE)

```




## What distance?
Here we are testing at what radius from which we create the plant composition do we see the STRONGEST correlation between plant community composition and microbial community. We will do this with both bacteria and fungi. 


```{r mantel_correlogram_all_samples__bacteria}
print("For BACTERIA") # just so we know what the output is for in the HTML file

### These matrices only need to be run once because they do not depend on distance

## MICROBIAL MATRIX
# Clean up the Bray-Curtis matrix for bacterial communities
load("03_Processed_data/16S_glass_slides.rda")
M.microbial <- q1.bray.dist.df %>% 
  select(starts_with("LO")) %>% 
  rename(LOT1A3 = LOT1A3_dup_new_protocol) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  column_to_rownames("SampleID") %>% as.matrix


## DISTANCE MATRIX
# Load site coordinates and use to make dataframe of coordinates for each sample
# site.coords <- as.data.frame(fread("03_Processed_data/site_locs_coordinates.tsv")) %>%
#   rename(Site = V1)
# 
# M.distance <- data.frame("Sample" = row.names(M.microbial)) %>% # start with same names in same order
#   mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>%  # get site ID
#   left_join(site.coords) %>% # join with our coordinates
#   column_to_rownames("Sample") %>%
#   select(x,y) %>% dist(., method = "euclidean") %>% # create Euclidean distance matrix
#   as.matrix


### Here we start the loop through our difference distances

## PLANT MATRIX
## reading in plant composition bc and make df of plant composition of each sample
plant.list <- list.files("03_Processed_data/map/", pattern = "_meter_composition_df.tsv")
r.list <- list()
p.list <- list()
radius.list <- list()

## The radius at which you measured plant composition
radius <- c(0.1, 0.5, 0.75, 1, 1.25, 1.5, 2, 2.5, 3, 3.5, 4) # COPY THIS FROM THE -- MAP_make_plant_composition_spatial_polygon_file.R

for (i in 1:length(plant.list)) {
  plants <- fread(paste0("03_Processed_data/map/", plant.list[[i]]), data.table = FALSE) %>% 
  rename(Site = V1)

M.plants <- data.frame("Sample" = row.names(M.microbial)) %>% 
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>% 
  left_join(plants) %>% 
  column_to_rownames("Sample") %>% 
  select(!(Site)) %>% 
  vegdist() %>% as.matrix

## mantel time!!
mantel.output <- vegan::mantel(M.microbial, M.plants, permutations = 10000, method = "pearson")


r.list[[i]] <- mantel.output$statistic
p.list[[i]] <- mantel.output$signif
}


mantel.corr <- data.frame("r" = unlist(r.list), 
                          "p" = unlist(p.list), 
                          "radius" = radius) %>% 
  mutate(signif = ifelse(p < 0.05, "yes", "no"))

ggplot(data = mantel.corr, aes(x = radius, y = r)) + 
  geom_line() + 
  geom_point(color = "white", size = 5) +
  geom_point(aes(shape = signif), 
                 colour = "black", size = 5) +
  scale_shape_manual(values=c(1,19) , guide = "none") + 
  theme_test() + 
  ggtitle("All Samples, Bacteria")
```


```{r mantel_correlogram_subset_samples__bacteria}
print("For BACTERIA") # just so we know what the output is for in the HTML file

### These matrices only need to be run once because they do not depend on distance

## MICROBIAL MATRIX
# Clean up the Bray-Curtis matrix for bacterial communities
load("03_Processed_data/16S_glass_slides.rda")
M.microbial <- q1.bray.dist.df %>%
  select(starts_with("LO") & contains(c("A2", "A7", "A8",
                                        "B1", "B2", "B7", "B8",
                                        "C1", "C2", "C7", "C8"))) %>% # A1 and B1 overlap so one has to be removed
  rownames_to_column("SampleID") %>%
  filter(grepl("LO", SampleID),
         grepl("A2|A7|A8|B1|B2|B7|B8|C1|C2|C7|C8", SampleID)) %>% # A1 and B1 overlap so one has to be removed
  column_to_rownames("SampleID") %>% as.matrix



## DISTANCE MATRIX
# Load site coordinates and use to make dataframe of coordinates for each sample
site.coords <- as.data.frame(fread("03_Processed_data/site_locs_coordinates.tsv")) %>%
  rename(Site = V1)

M.distance <- data.frame("Sample" = row.names(M.microbial)) %>% # start with same names in same order
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>%  # get site ID
  left_join(site.coords) %>% # join with our coordinates
  column_to_rownames("Sample") %>%
  select(x,y) %>% dist(., method = "euclidean") %>% # create Euclidean distance matrix
  as.matrix


## Simple Mantel test to check if distance and bacterial composition in these subsetted samples are correlated
vegan::mantel(M.microbial, M.distance, permutations = 10000, method = "pearson")
# yes, they are so we must do a partial mantel below



### Here we start the loop through our difference distances

## PLANT MATRIX
## reading in plant composition bc and make df of plant composition of each sample
plant.list <- list.files("03_Processed_data/map/", pattern = "_meter_composition_df.tsv")
r.list <- list()
p.list <- list()
radius.list <- list()

## The radius at which you measured plant composition
radius <- c(0.1, 0.5, 0.75, 1, 1.25, 1.5, 2, 2.5, 3, 3.5, 4) # COPY THIS FROM THE -- MAP_make_plant_composition_spatial_polygon_file.R


for (i in 1:length(plant.list)) {
  plants <- fread(paste0("03_Processed_data/map/", plant.list[[i]]), data.table = FALSE) %>% 
  rename(Site = V1)

M.plants <- data.frame("Sample" = row.names(M.microbial)) %>% 
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>% 
  left_join(plants) %>% 
  column_to_rownames("Sample") %>% 
  select(!(Site)) %>% 
  vegdist() %>% as.matrix


mantel.output <- partial.mantel.test(M.microbial, M.distance, M.plants, resamp = 10000, method = "pearson", quiet = TRUE)


r.list[[i]] <- mantel.output$MantelR[["r13.2"]]
p.list[[i]] <- mantel.output$p[[5]]


}


mantel.corr <- data.frame("r" = unlist(r.list), 
                          "p" = unlist(p.list), 
                          "radius" = radius) %>% 
  mutate(signif = ifelse(p < 0.05, "yes", "no"))

ggplot(data = mantel.corr, aes(x = radius, y = r)) + 
  geom_line() + 
  geom_point(color = "white", size = 5) +
  geom_point(aes(shape = signif), 
                 colour = "black", size = 5) +
  scale_shape_manual(values=c(1,19) , guide = "none") + 
  theme_test() + 
  ggtitle("No Overlapping Samples, Bacteria")
```




```{r mantel_correlogram_all_samples__fungi}
print("For FUNGI") # just so we know what the output is for in the HTML file

### These matrices only need to be run once because they do not depend on distance

## MICROBIAL MATRIX
# Clean up the Bray-Curtis matrix for bacterial communities
load("03_Processed_data/ITS_glass_slides.rda")
M.microbial <- q1.ITS.bray.dist.df %>% 
  select(starts_with("LO")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  column_to_rownames("SampleID") %>% as.matrix


## DISTANCE MATRIX
# Load site coordinates and use to make dataframe of coordinates for each sample
site.coords <- as.data.frame(fread("03_Processed_data/site_locs_coordinates.tsv")) %>% 
  rename(Site = V1)

M.distance <- data.frame("Sample" = row.names(M.microbial)) %>% # start with same names in same order
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>%  # get site ID
  left_join(site.coords) %>% # join with our coordinates
  column_to_rownames("Sample") %>% 
  select(x,y) %>% dist(., method = "euclidean") %>% # create Euclidean distance matrix
  as.matrix


### Here we start the loop through our difference distances

## PLANT MATRIX
## reading in plant composition bc and make df of plant composition of each sample
plant.list <- list.files("03_Processed_data/map/", pattern = "_meter_composition_df.tsv")
r.list <- list()
p.list <- list()
radius.list <- list()

## The radius at which you measured plant composition
radius <- c(0.1, 0.5, 0.75, 1, 1.25, 1.5, 2, 2.5, 3, 3.5, 4) # COPY THIS FROM THE -- MAP_make_plant_composition_spatial_polygon_file.R

for (i in 1:length(plant.list)) {
  plants <- fread(paste0("03_Processed_data/map/", plant.list[[i]]), data.table = FALSE) %>% 
  rename(Site = V1)

M.plants <- data.frame("Sample" = row.names(M.microbial)) %>% 
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>% 
  left_join(plants) %>% 
  column_to_rownames("Sample") %>% 
  select(!(Site)) %>% 
  vegdist() %>% as.matrix

## Partial mantel time!!
mantel.output <- partial.mantel.test(M.microbial, M.distance, M.plants, resamp = 10000, method = "pearson", quiet = TRUE)

r.list[[i]] <- mantel.output$MantelR[["r13.2"]]
p.list[[i]] <- mantel.output$p[[5]]
}

mantel.corr <- data.frame("r" = unlist(r.list), 
                          "p" = unlist(p.list), 
                          "radius" = radius) %>% 
  mutate(signif = ifelse(p < 0.05, "yes", "no"))

ggplot(data = mantel.corr, aes(x = radius, y = r)) + 
  geom_line() + 
  geom_point(color = "white", size = 5) +
  geom_point(aes(shape = signif), 
                 colour = "black", size = 5) +
  scale_shape_manual(values=c(19,1) , guide = "none") + 
  theme_test() + 
  ggtitle("All Samples, Fungi")
```



```{r mantel_correlogram_subset_samples__fungi}
print("For FUNGI") # just so we know what the output is for in the HTML file

### These matrices only need to be run once because they do not depend on distance

## MICROBIAL MATRIX
# Clean up the Bray-Curtis matrix for bacterial communities
load("03_Processed_data/ITS_glass_slides.rda")
M.microbial <- q1.ITS.bray.dist.df %>% 
  select(starts_with("LO") & contains(c("A2", "A7", "A8", 
                                        "B1", "B2", "B7", "B8", 
                                        "C1", "C2", "C7", "C8"))) %>% # A1 and B1 overlap so one has to be removed 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID), 
         grepl("A2|A7|A8|B1|B2|B7|B8|C1|C2|C7|C8", SampleID)) %>% # A1 and B1 overlap so one has to be removed 
  column_to_rownames("SampleID") %>% as.matrix


## DISTANCE MATRIX
# Load site coordinates and use to make dataframe of coordinates for each sample
site.coords <- as.data.frame(fread("03_Processed_data/site_locs_coordinates.tsv")) %>% 
  rename(Site = V1)

M.distance <- data.frame("Sample" = row.names(M.microbial)) %>% # start with same names in same order
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>%  # get site ID
  left_join(site.coords) %>% # join with our coordinates
  column_to_rownames("Sample") %>% 
  select(x,y) %>% dist(., method = "euclidean") %>% # create Euclidean distance matrix
  as.matrix


### Here we start the loop through our difference distances

## PLANT MATRIX
## reading in plant composition bc and make df of plant composition of each sample
plant.list <- list.files("03_Processed_data/map/", pattern = "_meter_composition_df.tsv")
r.list <- list()
p.list <- list()
radius.list <- list()

## The radius at which you measured plant composition
radius <- c(0.1, 0.5, 0.75, 1, 1.25, 1.5, 2, 2.5, 3, 3.5, 4) # COPY THIS FROM THE -- MAP_make_plant_composition_spatial_polygon_file.R

for (i in 1:length(plant.list)) {
  plants <- fread(paste0("03_Processed_data/map/", plant.list[[i]]), data.table = FALSE) %>% 
  rename(Site = V1)

M.plants <- data.frame("Sample" = row.names(M.microbial)) %>% 
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>% 
  left_join(plants) %>% 
  column_to_rownames("Sample") %>% 
  select(!(Site)) %>% 
  vegdist() %>% as.matrix

## Partial mantel time!!
mantel.output <- partial.mantel.test(M.microbial, M.distance, M.plants, resamp = 10000, method = "pearson", quiet = TRUE)

r.list[[i]] <- mantel.output$MantelR[["r13.2"]]
p.list[[i]] <- mantel.output$p[[5]]
}

mantel.corr <- data.frame("r" = unlist(r.list), 
                          "p" = unlist(p.list), 
                          "radius" = radius) %>% 
  mutate(signif = ifelse(p < 0.05, "yes", "no"))

ggplot(data = mantel.corr, aes(x = radius, y = r)) + 
  geom_line() + 
  geom_point(color = "white", size = 5) +
  geom_point(aes(shape = signif), 
                 colour = "black", size = 5) +
  scale_shape_manual(values=c(19,1) , guide = "none") + 
  theme_test() + 
  ggtitle("No Overlapping Samples, Fungi")
```
