---
title: "Question_2__drivers_of_heterogeneity"
author: "Kendra Walters"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
.main-container { max-width: 1000px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

---

# Purpose
To test whether plant composition is the main driver of heterogeneity (or, other option, spatial distance which may stand in for other factor(s)) AND, if it is, then to test at what radius do we see the strongest impact of plant composition. 

```{r setup, message=FALSE}
library(knitr)
library(tidyverse)
library(data.table)
library(kableExtra)
require(ggplot2)
require(gridExtra)
library(car)
require(vegan)
require(magrittr)
library(indicspecies)
require(pairwiseAdonis)

knitr::opts_knit$set(root.dir = "../")

knitr::opts_chunk$set(comment="", cache=T, warning = F, message = F, fig.path = "05_output_figures/")

```

## Distance v. plants
It's like aliens versus plants but much less interactive. Basically, we are doing a partial mantel here to see if geographic distance and/or plant composition is driving community composition in the bacterial and fungal dispersal communities found on the glass slides. For this analysis, we are arbitrarily using the plant composition within 1 meter of each site. We could have chosen a different radius here, but this one seemed OK.

```{r distance_v_plants__bacteria}
print("For BACTERIA") # just so we know what the output is for in the HTML file


## MICROBIAL MATRIX
# Clean up the Bray-Curtis matrix for bacterial communities
load("03_Processed_data/16S_glass_slides.rda")
M.microbial <- q1.bray.dist.df %>% 
  select(starts_with("LO")) %>% 
  rename(LOT1A3 = LOT1A3_dup_new_protocol) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  column_to_rownames("SampleID") %>% as.matrix


## DISTANCE MATRIX
# Load site coordinates and use to make dataframe of coordinates for each sample
site.coords <- as.data.frame(fread("03_Processed_data/site_locs_coordinates.tsv")) %>% 
  rename(Site = V1)

M.distance <- data.frame("Sample" = row.names(M.microbial)) %>% # start with same names in same order
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>%  # get site ID
  left_join(site.coords) %>% # join with our coordinates
  column_to_rownames("Sample") %>% 
  select(x,y) %>% dist(., method = "euclidean") %>% # create Euclidean distance matrix
  as.matrix


## PLANT MATRIX
## reading in plant composition bc and make df of plant composition of each sample
plants <- fread("grass_shrub_1_meter_composition_df.tsv", data.table = FALSE)

sample.plants <- data.frame("Sample" = row.names(bc))
sample.plants$Site <- str_extract(sample.plants$Sample, "[ABC][1-8]")
sample.plants <- merge(sample.plants, plants, by.x = "Site", by.y = "V1")

plants <- fread("03_Processed_data/map/grass_shrub_1_meter_composition_df.tsv", data.table = FALSE) %>% 
  rename(Site = V1)

M.plants <- data.frame("Sample" = row.names(M.microbial)) %>% 
  mutate(Site = str_extract(Sample, "[ABC][1-8]")) %>% 
  left_join(plants) %>% 
  column_to_rownames("Sample") %>% 
  select(!(Site)) %>% 
  vegdist() %>% as.matrix



## Partial mantel time!!
partial.mantel.test(M.microbial, M.distance, M.plants, resamp = 1000, method = "pearson", quiet = FALSE)

```



```{r distance_v_plants__fungi}
print("For FUNGI") # just so we know what the output is for in the HTML file
```
