---
title: "Question_1__characterizing_dispersal"
author: "Kendra E. Walters"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
.main-container { max-width: 1000px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

# Purpose

To answer our first overarching question for the landscape dispersal project: Is the rate and composition of microbial dispersal into the soil surface spatially heterogeneous? 

Because vegetation appears to be a primary route through which bacteria disperse, we hypothesized that the taxa dispersing and their dispersal rate onto the soil surface will differ between two ecosystems. In fact, we know that bacterial abundance and composition in one grassland and CSS plant litter differs.


```{r setup, message=FALSE}
library(knitr)
library(tidyverse)
library(data.table)
library(kableExtra)
require(ggplot2)
require(gridExtra)
library(car)
require(vegan)
require(magrittr)
library(indicspecies)
require(pairwiseAdonis)
require(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")

knitr::opts_knit$set(root.dir = "../")

knitr::opts_chunk$set(comment="", cache=T, warning = F, message = F, fig.path = "05_output_figures/")


```

# Abundance on glass slides

First, we'll look at the abundances on the glass slides over time to see if they are acculumulating or at equilibrium. 

```{r boxplot_glass_slide_abundance_by_ecosystem}
# Load the data
load("03_Processed_data/flow_cytometry_cleaned.rda")
load("03_Processed_data/metadata_cleaned.rda")


# Left join fights me unless both columns are the same data type
metadata.together$`Random_#_Flow` <- as.character(metadata.together$`Random_#_Flow`)


# Set up a key to change time point into # of days
days.key <- c("0" = 0, "1" = 12, "2" = 26, "3" = 47, "4" = 54)


# Clean up our data to just be what we need for this analysis 
cell.counts.use <- cleaned.cell.counts %>% left_join(metadata.together) %>% # combine with metadata using Random_#_Flow
  mutate(Treatment = ifelse(is.na(Treatment), "Death", Treatment)) %>% # add metadata for T0 samples that weren't included in the overall metadata file
  mutate(Time_Point = ifelse(is.na(Time_Point), 0, Time_Point)) %>% 
  mutate(Ecosystem = ifelse(grepl("^g", `Random_#_Flow`), "Grassland",
                            ifelse(grepl("^s", `Random_#_Flow`), "Shrubland", Ecosystem))) %>%
  select(number_cells_on_glass_slide_cleaned, Treatment, Ecosystem, Time_Point) %>% # reduce to only columns we need
  filter(Treatment != "Closed") %>% # remove this negative control (should have been removed prior to this script)
  mutate(Time_as_days = days.key[as.character(Time_Point)]) %>% # make column that lists # of days
  mutate(cells_per_sq_cm = number_cells_on_glass_slide_cleaned / (7.5 * 2.5))


# Plot this as a boxplot to give a general sense of what the abundance on the glass slides are doing
ggplot(data = cell.counts.use %>% filter(Treatment == "Open")) +
  geom_boxplot(aes(x = as.factor(Time_Point), y = cells_per_sq_cm)) + 
  facet_wrap(vars(Ecosystem))
```

# Immigration rate calculations assuming population growth

Okay, seems like the immigration rate is faster than the death rate for the first three timepoints, and then a plateau is potentially being reached. We can model this increase and plateau using the integral of our population "growth" equation (with just immigration and death). 

dn/dt = i - d*n(t) where n = abundance on glass slides, i = immigration rate, and d = death rate

So the integral is (when forced to go through the origin): 
n(t) = i/d * (1 - e^(-d*t))

```{r scatterplot_abundance_glass_slides_by_ecosystem}
# plotting our data with our fancy integral equation going through the origin
ggplot(data = cell.counts.use %>% filter(Treatment == "Open") %>% 
  mutate(Time_as_days= as.integer(Time_as_days)), 
       aes(x = Time_as_days, y = cells_per_sq_cm)) +
  geom_point() + 
  geom_smooth(method = "nls", formula = "y ~  (Im / d) * (1 - exp(-d * x))", se=F, 
              method.args = list(start=c(Im=2e5, d=0.034))) +
  facet_wrap(vars(Ecosystem)) 


# Separate into our two ecosystems
grass.cells <- cell.counts.use %>% filter(Ecosystem == "Grassland")
shrub.cells <- cell.counts.use %>% filter(Ecosystem == "Shrubland")


# Calculating the death rate and immigration rate coefficients
(nls.shrub <- nls(cells_per_sq_cm ~  Im/d * (1 - exp(-d * Time_as_days)), 
               data = shrub.cells[shrub.cells$Treatment == "Open", ],
               start = list(Im = 2e5, d = 0.034)) %>% summary) 
(nls.grass <- nls(cells_per_sq_cm ~  Im/d * (1 - exp(-d * Time_as_days)), 
               data = grass.cells[grass.cells$Treatment == "Open", ],
               start = list(Im = 2e5, d = 0.034)) %>% summary)

# note: use summary(model) to look at standard error, which is HUGE for the death rate and fairly large for the immigration rate
```

But we don't actually need to have the model guess what the death rate is because we measured death rate with our death rate glass slides. So let's plot that data now and calculate our death rate for the grassland and shrubland. 

```{r scatterplot_death_rate_by_ecosystem}
# Calculating death rate for the grassland death slides
intercept.grass <- grass.cells %>% 
  filter(Treatment == "Death", 
         Time_as_days == 0) %>% 
  summarize(mean = mean(log(cells_per_sq_cm))) %>% pull(mean)

r.grass <- lm(I(log(cells_per_sq_cm) - intercept.grass) ~ Time_as_days + 0, 
               grass.cells %>% filter(Treatment == "Death")) %>% coef*-1 # THIS IS THE DEATH RATE!! (we do negative because this is actually describing the continuous rate at which cells survive, not the continuous rate at which cells die)
print(paste0("Continuous death rate for grassland cells is ", round(r.grass, 4), "."))
print(paste0("Daily death rate for grassland cells is " , round(-100*(1 - exp(r.grass * 1)), 2), "% per day."))
  

# Calculating death rate for the grassland death slides
intercept.shrub <- shrub.cells %>% 
  filter(Treatment == "Death", 
         Time_as_days == 0) %>% 
  summarize(mean = mean(log(cells_per_sq_cm))) %>% pull(mean)

r.shrub <- lm(I(log(cells_per_sq_cm) - intercept.shrub) ~ Time_as_days + 0, 
               shrub.cells %>% filter(Treatment == "Death")) %>% coef*-1 # THIS IS THE DEATH RATE!! (we do negative because this is actually describing the continuous rate at which cells survive, not the continuous rate at which cells die)
print(paste0("Continuous death rate for shrubland cells is ", round(r.shrub, 4), "."))
print(paste0("Daily death rate for grassland cells is " , round(-100*(1 - exp(r.shrub * 1)), 2), "% per day."))


# creating a mini data frame to give our parameters for the linear models (since we are picky about our intercepts here)
lm.parameters.death <- data.frame("Intercept" = c(intercept.grass, intercept.shrub), 
                                  "Slope" = c(r.grass[[1]], r.shrub[[1]]), 
                                  "Ecosystem" = c("Grassland", "Shrubland"))


# plotting our death rate data
ggplot(data = cell.counts.use %>% filter(Treatment == "Death") %>% 
  mutate(Time_as_days= as.integer(Time_as_days)), 
       aes(x = Time_as_days, y = log(cells_per_sq_cm))) +
  geom_point() + 
  geom_abline(data = lm.parameters.death, aes(intercept = Intercept, slope = -Slope)) + 
  facet_wrap(vars(Ecosystem))

```

Sweet! Now we can use these death rates to input into our model so we are only asking it to find the estimate for the immigration rate. Note how much smaller the standard errors are for these immigration rates! 

```{r scatterplot_abundance_glass_slides_with_measured_death_rate_by_ecosystem}
# Calculating the death rate and immigration rate coefficients WITH measured death rates
(nls.shrub <- nls(cells_per_sq_cm ~  
                    Im/0.02372169 * (1 - exp(-0.02372169 * Time_as_days)), 
               data = shrub.cells[shrub.cells$Treatment == "Open", ],
               start = list(Im = 2e5)) %>% summary) 
(nls.grass <- nls(cells_per_sq_cm ~ 
                    Im/0.03106574 * (1 - exp(-0.03106574 * Time_as_days)), 
               data = grass.cells[grass.cells$Treatment == "Open", ],
               start = list(Im = 2e5)) %>% summary)

nls.grass$parameters[[1]]

# Calculate % of typical litterbag community
fread("01_Raw_data/BacterialAbundanceLomaRidgeLitterbagsFromClaudia.txt", 
                    data.table = FALSE) %>% 
  filter(timepoint != 0, 
         moTreatment == "ambient") %>% 
  mutate(`BACavgcount/g dry weight` = as.numeric(`BACavgcount/g dry weight`), 
         AbundancePerBag = `DryMass (g)` * `BACavgcount/g dry weight`, 
         AbundancePer_cm2 = AbundancePerBag / (13 * 14)) %>% # assuming a 15x15cm bag with 1 cm on each of the 3 sealed sides (1 side was created by folding over the 30cm x 15cm nylon sheet)
  group_by(LitterOrigin) %>% 
  summarize(mean = mean(AbundancePer_cm2)) %>% 
  left_join(data.frame("LitterOrigin" = c("grass", "shrub"), 
                       "immigration_rate" = c(nls.grass$parameters[[1]], nls.shrub$parameters[[1]]))) %>% 
  mutate(immigration_percent = (immigration_rate / mean)*100) %>% 
  print(.)


# Graph the new curves that we made
A <- ggplot(data = grass.cells %>% filter(Treatment == "Open") %>% 
  mutate(Time_as_days= as.integer(Time_as_days)), 
       aes(x = Time_as_days, y = cells_per_sq_cm)) +
  geom_point() + 
  geom_smooth(method = "nls", formula = "y ~  (Im / 0.03106574) * (1 - exp(-0.03106574 * x))", se=F, 
              method.args = list(start=c(Im=2e5))) +
  ggtitle("Grassland")

B <- ggplot(data = shrub.cells %>% filter(Treatment == "Open") %>% 
  mutate(Time_as_days= as.integer(Time_as_days)), 
       aes(x = Time_as_days, y = cells_per_sq_cm)) +
  geom_point() + 
  geom_smooth(method = "nls", formula = "y ~  (Im / 0.02372169) * (1 - exp(-0.02372169 * x))", se=F, 
              method.args = list(start=c(Im=2e5))) +
  ggtitle("Shrubland")

grid.arrange(A,B, nrow = 1)

```

# Immigration rate calculations assuming equilibrium
Looking at the first graph of abundance on the glass slides above, it seems that, on and after the third time point, the abundance on the glass slides is at equilibrium. Actually, for the grassland, it looks like abundance dips a little, either driven by higher death rate or lower immigration rate, so we can just use the third timepoint to calculate our immigration rate at equilibrium. 


```{r immigration_rate_at_equilibrium}
# Calculate immigration rate for the grassland 
imm.grass <- grass.cells %>% filter(Treatment == "Open", Time_Point %in% c("3")) %>% 
  summarize(mean = mean(cells_per_sq_cm)) * r.grass

print(paste0("The immigration rate for the grassland is ", imm.grass, " cells / cm^2 / day."))

# Calculate immigration rate for the shrubland 
imm.shrub <- shrub.cells %>% filter(Treatment == "Open", Time_Point %in% c("3")) %>% 
  summarize(mean = mean(cells_per_sq_cm)) * r.shrub

print(paste0("The immigration rate for the shrubland is ", imm.shrub, " cells / cm^2 / day."))
```


# Death rates by ecosystem
One potentially interesting question is if death rates (for bacteria) differ by ecosystem. We can answer this question by running an ANCOVA on the abundance in the death slides, to see if there is an interaction between time and ecosystem. 
```{r death_rates_by_ecosystem}
mod <- cell.counts.use %>% filter(Treatment == "Death") %>% 
  mutate(log_abundance_minus_intercept = ifelse(Ecosystem == "Shrubland", 
                                            log(cells_per_sq_cm) - intercept.shrub, 
                              ifelse(Ecosystem == "Grassland", 
                                     log(cells_per_sq_cm) - intercept.grass, NA))) %>% 
  lm(formula = log_abundance_minus_intercept ~ Time_as_days:Ecosystem + Time_as_days + 0, data = .) %>% # independent of order
  Anova(type = "II") # MARGINALLY SIGNIFICANT

mod

print(paste0("The r2 for time is: ", 
             mod$`Sum Sq`[1] / (mod$`Sum Sq`[1] + mod$`Sum Sq`[2] + mod$`Sum Sq`[3])))
print(paste0("The r2 for time-by-ecosystem interaction is: ", 
             mod$`Sum Sq`[2] / (mod$`Sum Sq`[1] + mod$`Sum Sq`[2] + mod$`Sum Sq`[3])))
print(paste0("The r2 for the residuals is: ", 
             mod$`Sum Sq`[3] / (mod$`Sum Sq`[1] + mod$`Sum Sq`[2] + mod$`Sum Sq`[3])))
```

# Taxa dying at different rates? 
Another potentially interesting question is if taxa on the death rate slides die at different rates. We can answer this question by running PERMANOVA on the death rate samples to see if time influences either 16S or ITS composition. Because there are so few samples that sequenced well after timepoint 1, we are grouping 2-4 to run the PERMANOVA. 
```{r NMDS_taxa_specific_death_rates_bacteria}
# Load data
load("03_Processed_data/16S_glass_slides.rda")

tmp <- q1.bray.dist.df %>%  select(starts_with("LD")) %>% 
                                       select(!("LDT0R5G_oops")) %>% 
                                       rownames_to_column("SampleID") %>% 
                                       filter(grepl("LD", SampleID)) %>% 
                                       filter(!(SampleID == "LDT0R5G_oops")) %>% 
                                       column_to_rownames("SampleID")


# Create our visualization, just for fun
q1.NMDS <- metaMDS(as.dist(as.matrix(q1.bray.dist.df %>% 
                                       select(starts_with("LD")) %>% 
                                       select(!("LDT0R5G_oops")) %>% 
                                       rownames_to_column("SampleID") %>% 
                                       filter(grepl("LD", SampleID)) %>% 
                                       filter(!(SampleID == "LDT0R5G_oops")) %>% 
                                       column_to_rownames("SampleID"))), 
                   autotransform = FALSE, k = 2, trymax = 200, trace = 0)
q1.NMDS.meta <- data.frame(q1.NMDS$points[,1:2]) %>% 
  rownames_to_column("SampleID") %>% 
  mutate(TimePoint = str_match(SampleID, "T([0-4])")[,2]) %>% 
  mutate(Ecosystem = str_match(SampleID, "R[0-9][A-C]?([GS])")[,2], 
         Ecosystem = ifelse(Ecosystem == "G", "Grassland", "Shrubland")) %>% 
  mutate(TimePoint_grouped = ifelse(grepl("0|1", TimePoint), TimePoint, "2_through_4"))

ggplot(data = q1.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(TimePoint), shape = as.factor(Ecosystem)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Bacteria, death slides over time")

# Do the PERMANOVA that will test our hypothesis
q1.bray.dist.df.death <- as.data.frame(as.matrix(q1.bray.dist)) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LD", SampleID)) %>% 
  filter(!(SampleID == "LDT0R5G_oops")) %>% 
  mutate(SampleID = str_remove(SampleID, "_dup_new_protocol")) %>% 
  column_to_rownames("SampleID") %>% 
  select(starts_with("LD")) %>%
  select(!("LDT0R5G_oops")) %>% 
  rename(LDT1R5BG = LDT1R5BG_dup_new_protocol)

      
adonis(q1.bray.dist.df.death ~ TimePoint_grouped, permutations = 10000, 
       data = q1.NMDS.meta, na.rm= TRUE)

adonis(q1.bray.dist.df.death ~ TimePoint_grouped * as.factor(Ecosystem), permutations = 10000, 
       data = q1.NMDS.meta, na.rm= TRUE)
```


```{r NMDS_taxa_specific_death_rates_fungi}
# Load data
load("03_Processed_data/ITS_glass_slides.rda")

# Create our visualization, just for fun
q1.ITS.NMDS <- metaMDS(as.dist(as.matrix(q1.ITS.bray.dist.df %>% 
                                           select(starts_with("LD")) %>% 
                                           rownames_to_column("SampleID") %>% 
                                           filter(grepl("LD", SampleID)) %>% 
                                           column_to_rownames("SampleID"))), 
                       autotransform = FALSE, k = 2, trymax = 200, trace = 0)

q1.ITS.NMDS.meta <- data.frame(q1.ITS.NMDS$points[,1:2]) %>% 
  rownames_to_column("SampleID") %>% 
  mutate(TimePoint = str_match(SampleID, "T([0-4])")[,2]) %>% 
  mutate(Ecosystem = str_match(SampleID, "R[0-9][A-C]?([GS])")[,2], 
         Ecosystem = ifelse(Ecosystem == "G", "Grassland", "Shrubland")) %>% 
  mutate(TimePoint_grouped = ifelse(grepl("0|1", TimePoint), TimePoint, "2_through_4"))

ggplot(data = q1.ITS.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(TimePoint), shape = as.factor(Ecosystem)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Fungi, death slides over time")

# Do the PERMANOVA that will test our hypothesis
q1.bray.dist.df.death <- q1.ITS.bray.dist.df %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LD", SampleID)) %>% 
  column_to_rownames("SampleID") %>% 
  select(starts_with("LD"))


adonis(q1.bray.dist.df.death ~ TimePoint_grouped, permutations = 10000, 
       data = q1.ITS.NMDS.meta, na.rm= TRUE)

adonis(q1.bray.dist.df.death ~ TimePoint_grouped * as.factor(Ecosystem), permutations = 10000, 
       data = q1.ITS.NMDS.meta, na.rm= TRUE)

```


# Dispersal communities by ecosystem
A key question of this study is - Do different ecosystems (plant communities) have different dispersal communities? To answer this question, we will use a PERMANOVA to see if different ecosystems show different bacterial and fungal communities on the glass slides. However, because variance seems to change through time along with the community composition, we will use a PERMDISP to see if variance could be driving the significant PERMANOVA too. And then, because the pattern over time is clearer with the fungi, we will look to see if one or a few taxa are driving this relationship. 

```{r NMDS_dispersal_communities_by_ecosystem_bacteria}
# Load data
load("03_Processed_data/16S_glass_slides.rda")

# Subset our Bray-Curtis distance matrix to include only samples we want to plot and test
q1.bray.dist.df.dispersal <- q1.bray.dist.df %>% 
  select(contains("LO")) %>% 
  rename(LOT1A3 = LOT1A3_dup_new_protocol) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  column_to_rownames("SampleID")

# Do the NMDS
q1.dispersal.NMDS <- metaMDS(as.dist(as.matrix(q1.bray.dist.df.dispersal)), 
                             autotransform = FALSE, k = 2, trymax = 200, 
                             trace = 0)

# Merge with our metatdata and create a new column to specify what we want to look at
q1.NMDS.meta <- data.frame(q1.dispersal.NMDS$points[,1:2]) %>% 
  rownames_to_column("New_Sample_ID") %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Time_Point, Ecosystem))

# Create our visualization of dispersal slides + sources
ggplot(data = q1.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(Ecosystem)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Bacteria, glass slides by ecosystem")

ggplot(data = q1.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, shape = as.factor(Ecosystem), color = as.factor(Time_Point)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Bacteria, glass slides by ecosystem AND time")

# Do the staaaats!!
print("These stats are for the BACTERIA")

adonis(q1.bray.dist.df.dispersal ~ Ecosystem, permutations = 10000, 
       data = q1.NMDS.meta, na.rm= TRUE)

adonis(q1.bray.dist.df.dispersal ~ Ecosystem * Time_Point, permutations = 10000, 
       data = q1.NMDS.meta, na.rm= TRUE) # interaction is not significant 




# See if variance may be partially driving the significant time effect on composition
print("To see if TIME POINT has different variances and may be driving significant time e effect on composition")
betadisper(as.dist(as.matrix(q1.bray.dist.df.dispersal)), q1.NMDS.meta$Time_Point, type = "centroid") %>% # no interaction in the PERMANOVA, so we can test each factor independently here
  anova()
```

```{r NMDS_dispersal_communities_by_ecosystem_fungi}
# Load data
load("03_Processed_data/ITS_glass_slides.rda")

# Subset our Bray-Curtis distance matrix to include only samples we want to plot and test
q1.ITS.bray.dist.df.dispersal <- q1.ITS.bray.dist.df %>% 
  select(contains("LO")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  column_to_rownames("SampleID")

# Do the NMDS
q1.ITS.dispersal.NMDS <- metaMDS(as.dist(as.matrix(q1.ITS.bray.dist.df.dispersal)), 
                             autotransform = FALSE, k = 2, trymax = 200, trace = 0)

# Merge with our metatdata and create a new column to specify what we want to look at
q1.ITS.NMDS.meta <- data.frame(q1.ITS.dispersal.NMDS$points[,1:2]) %>% 
  rownames_to_column("New_Sample_ID") %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Time_Point, Ecosystem))

# Create our visualization of dispersal slides + sources
ggplot(data = q1.ITS.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(Ecosystem)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Fungi, glass slides by ecosystem")


ggplot(data = q1.ITS.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, shape = as.factor(Ecosystem), color = as.factor(Time_Point)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Fungi, glass slides by ecosystem AND time")

# Do the staaaats!!
print("These stats are for the FUNGI")

adonis(q1.ITS.bray.dist.df.dispersal ~ Ecosystem, permutations = 10000, 
       data = q1.ITS.NMDS.meta, na.rm= TRUE)

adonis(q1.ITS.bray.dist.df.dispersal ~ Ecosystem * Time_Point, permutations = 10000, 
       data = q1.ITS.NMDS.meta, na.rm= TRUE)

# See if variance may be partially driving the significant time effect on composition
print("To see if TIME POINT has different variances and may be driving significant time e effect on composition")
betadisper(as.dist(as.matrix(q1.ITS.bray.dist.df.dispersal)), q1.ITS.NMDS.meta$Time_Point, type = "centroid") %>% # no interaction in the PERMANOVA, so we can test each factor independently here
  anova()

# Because the influence of time is more clear with the fungi, let's look to see if a single taxon is driving this


# # Option one: import into PRIMER, and plot species that are correlated (above a certain r) with community variation as vectors on their ordinations
# # These scripts output the files needed to import into PRIMER
# q1.ITS.rarefied %>% 
#   rownames_to_column("SampleID") %>% 
#   filter(grepl("LO", SampleID)) %>% 
#   write_tsv("~/Google Drive/Dispersal_Ch_3/03_GitHub_Data_and_Scripts/landscape_dispersal/03_Processed_data/ITS_glass_slides_q1__rarefied_filtered_for_primer.tsv", col_names = TRUE)
# 
# q1.ITS.NMDS.meta %>% 
#   write_tsv("~/Google Drive/Dispersal_Ch_3/03_GitHub_Data_and_Scripts/landscape_dispersal/03_Processed_data/ITS_glass_slides_q1__NMDS_meta_filtered_for_primer.tsv", col_names = TRUE)
# 
# # From PRIMER - NMDS, correlated at 0.68 (Spearman), seems to drive strongly towards T2-4
# ITS.species <- fread(input = "01_Raw_data/03_Microbial_Community_Composition/glass_slides__ITS/13_taxonomy_filtered.tsv") %>% 
#   filter(grepl("6d360", `#OTUID`)) %>% 
#   select(taxonomy) %>% print #k__Fungi;p__Ascomycota;c__Dothideomycetes;o__Pleosporales;f__Didymosphaeriaceae;g__Paraconiothyrium;s__unidentified


## Option 2 (my favorite)
## try with an indicator species analysis (between T1 and T2-4, because that's all that's significant)
key <- c("1" = "1", "2" = "not_1", "3" = "not_1", "4" = "not_1")
#want: rows = sites
indic.time <- multipatt(q1.ITS.rarefied %>% 
                   rownames_to_column("SampleID") %>% 
                   filter(grepl("LO", SampleID)) %>% 
                   column_to_rownames("SampleID"), 
                 key[q1.ITS.NMDS.meta$Time_Point], duleg = TRUE)
summary(indic.time)

ITS.taxonomy <- fread(input = "01_Raw_data/03_Microbial_Community_Composition/glass_slides__ITS/13_taxonomy_filtered.tsv")

# Time Point 1, 5 top taxa
ITS.taxonomy %>% filter(grepl("6b550ad3889766092ea1d7be05293622", `#OTUID`)) %>% 
  select(taxonomy) %>% print # Truncatella_angustata
ITS.taxonomy %>% filter(grepl("3cc206a0424f7c33f75fe152cd42d757", `#OTUID`)) %>% 
  select(taxonomy) %>% print # f__Phaeosphaeriaceae
ITS.taxonomy %>% filter(grepl("adffc920dd86ce72a6385dd2063ee499", `#OTUID`)) %>% 
  select(taxonomy) %>% print # o__Dothideales
ITS.taxonomy %>% filter(grepl("dba6ff41ca6ac3dd820c1a20c67d1b18", `#OTUID`)) %>% 
  select(taxonomy) %>% print # s__Teichospora_mariae
ITS.taxonomy %>% filter(grepl("fd3bb0f93dc5043995cc6df1698a570c", `#OTUID`)) %>% 
  select(taxonomy) %>% print # f__Lophiostomataceae


# Time Point 2/3/4, 5 top taxa
ITS.taxonomy %>% filter(grepl("6d360c9c285bc318a1e9cbfd2a85a057", `#OTUID`)) %>% 
  select(taxonomy) %>% print # g__Paraconiothyrium
ITS.taxonomy %>% filter(grepl("ec3d78df700e0518a81ed0843c9d1d1c", `#OTUID`)) %>% 
  select(taxonomy) %>% print # f__Didymellaceae
ITS.taxonomy %>% filter(grepl("0ec291fa8414302b18ab7bfb89c33209", `#OTUID`)) %>% 
  select(taxonomy) %>% print # g__Alternaria
ITS.taxonomy %>% filter(grepl("3a61b3e1ee7a53548fa1eb416917ab5b", `#OTUID`)) %>% 
  select(taxonomy) %>% print # g__Alternaria
ITS.taxonomy %>% filter(grepl("347b17563a35eda7d903ea07a289fce2", `#OTUID`)) %>% 
  select(taxonomy) %>% print # g__Alternaria


# ## Option 3: do a CAP analysis
# tmp <- 
# capscale(formula =  q1.ITS.rarefied %>% 
#                    rownames_to_column("SampleID") %>% 
#                    filter(grepl("LO", SampleID)) %>% 
#                    column_to_rownames("SampleID") ~ Time_Point, q1.ITS.NMDS.meta, 
#          dist = "bray")
# plot(tmp)

```

# Alpha-diveristy of dispersal community
```{r alpha_diversity_by_ecosystem__bacteria}
load("03_Processed_data/16S_glass_slides.rda")

# Calculate alpha diversity! and filter to just glass slides
alpha.16S <- data.frame("Shannon" = diversity(q1.rarefied, index = "shannon")) %>% 
  mutate(Richness = specnumber(q1.rarefied), 
         Simpson = diversity(q1.rarefied, index = "simpson")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  left_join(metadata.together %>% 
              select(New_Sample_ID, Time_Point, Transect, Ecosystem), 
            by = c("SampleID" = "New_Sample_ID"))

## Calculate average alpha-diversity overall and by ecosystem, testing significance
print(paste0("The average RICHNESS for BACTERIA is: ", round(mean(alpha.16S$Richness), 0), " OTUs per glass slide (rarefied to 971 sequences)."))

print("And averaged by ecosystem is:")
aggregate(alpha.16S$Richness, by = list(alpha.16S$Ecosystem), FUN = mean)

anova(lm(Richness ~ Ecosystem, data = alpha.16S))
```

```{r alpha_diversity_by_ecosystem__fungi}
load("03_Processed_data/ITS_glass_slides.rda")

# Calculate alpha diversity! and filter to just glass slides
alpha.ITS <- data.frame("Shannon" = diversity(q1.ITS.rarefied, index = "shannon")) %>% 
  mutate(Richness = specnumber(q1.ITS.rarefied), 
         Simpson = diversity(q1.ITS.rarefied, index = "simpson")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  left_join(metadata.together %>% 
            select(New_Sample_ID, Time_Point, Transect, Ecosystem), 
          by = c("SampleID" = "New_Sample_ID"))

## Calculate average alpha-diversity overall and by ecosystem, testing significance
print(paste0("The average RICHNESS for FUNGI is: ", round(mean(alpha.ITS$Richness), 0), " OTUs per glass slide (rarefied to 2400 sequences)."))

print("And averaged by ecosystem is:")
aggregate(alpha.ITS$Richness, by = list(alpha.ITS$Ecosystem), FUN = mean)

anova(lm(Richness ~ Ecosystem, data = alpha.ITS))
```

```{r alpha_diversity_by_ecosystem__comparing_fungi_AND_bacteria}
print("Now we are COMPARING bacteria and fungi (both rarefied to 971 here):")

load("03_Processed_data/16S_glass_slides.rda")
alpha.compare.16S <- data.frame("Shannon" = diversity(q1.rarefied, index = "shannon")) %>% 
  mutate(Richness = specnumber(q1.rarefied), 
         Simpson = diversity(q1.rarefied, index = "simpson")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  left_join(metadata.together %>% 
              select(New_Sample_ID, Time_Point, Transect, Ecosystem), 
            by = c("SampleID" = "New_Sample_ID")) %>% 
  mutate(Taxa = "Bacteria")


q1.ITS.rarefied.971 <- fread("03_Processed_data/ITS_glass_slides__rarefied_971_rounded.tsv", data.table = FALSE) %>% 
  column_to_rownames("V1")

alpha.compare <- data.frame("Shannon" = diversity(q1.ITS.rarefied.971, index = "shannon")) %>% 
  mutate(Richness = specnumber(q1.ITS.rarefied.971), 
         Simpson = diversity(q1.ITS.rarefied.971, index = "simpson")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  left_join(metadata.together %>% 
            select(New_Sample_ID, Time_Point, Transect, Ecosystem), 
          by = c("SampleID" = "New_Sample_ID")) %>% 
  mutate(Taxa = "Fungi") %>% 
  bind_rows(alpha.compare.16S)

print(paste0("The average taxa richness for bacteria and fungi are:"))
aggregate(alpha.compare$Richness, by = list(alpha.compare$Taxa), FUN = mean)
anova(lm(Richness ~ Taxa, data = alpha.compare))

mod <- anova(lm(Richness ~ Taxa, data = alpha.compare))
print(paste0("And the r2 for that ANOVA is: ", 
             mod$`Sum Sq`[1] / (mod$`Sum Sq`[1] + mod$`Sum Sq`[2])))
```

# Beta-diversity of dispersal community by ecosystem
We hypothesize that dispersal communities will show higher beta-diversity in the shrubland because the plant community is more heterogeneous in the shrubland. So let's test that with our good friend, PERMDISP!

```{r betadisper_bacteria}
# Load data
load("03_Processed_data/16S_glass_slides.rda")

# Subset our Bray-Curtis distance matrix to include only samples we want to plot and test
q1.bray.dist.df.dispersal <- q1.bray.dist.df %>% 
  select(contains("LO")) %>% 
  rename(LOT1A3 = LOT1A3_dup_new_protocol) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  column_to_rownames("SampleID")

q1.permdisp.meta <- data.frame("New_Sample_ID" = names(q1.bray.dist.df.dispersal)) %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Time_Point, Ecosystem))

# Test betadiversity
betadisper(as.dist(as.matrix(q1.bray.dist.df.dispersal)), 
           q1.permdisp.meta$Ecosystem, 
           type = c("centroid")) %>% anova
```


```{r betadisper_fungi}
# Load data
glass.slides.ITS <- "03_Processed_data/ITS_glass_slides.rda"
load(glass.slides.ITS)

# Subset our Bray-Curtis distance matrix to include only samples we want to plot and test
q1.ITS.bray.dist.df.dispersal <- q1.ITS.bray.dist.df %>% 
  select(contains("LO")) %>% 
  rownames_to_column("SampleID") %>% 
  filter(grepl("LO", SampleID)) %>% 
  column_to_rownames("SampleID")

q1.ITS.permdisp.meta <- data.frame("New_Sample_ID" = names(q1.ITS.bray.dist.df.dispersal)) %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Time_Point, Ecosystem))

# Test betadiversity
betadisper(as.dist(as.matrix(q1.ITS.bray.dist.df.dispersal)), 
           q1.ITS.permdisp.meta$Ecosystem, 
           type = c("centroid")) %T>% print %>% anova %>% print
```

# Dispersal sources - Air, Environmental Litter, and Soil
Let's check out how our dispersal slides compare to the dispersal sources! We'll look at this with an NMDS, and maybe a PERMANOVA, and then with SourceTracker to get % of source communities. 

```{r NMDS_dispersal_sources_bacteria}
# Load data
glass.slides.16S <- "03_Processed_data/16S_glass_slides.rda"
load(glass.slides.16S)

# Subset our Bray-Curtis distance matrix to include only samples we want to plot and test
q1.bray.dist.df.source <- q1.bray.dist.df %>% 
  select(!(contains("LD"))) %>% 
  rename(LOT1A3 = LOT1A3_dup_new_protocol) %>% 
  rownames_to_column("SampleID") %>% 
  filter(!(grepl("LD", SampleID))) %>% 
  mutate(SampleID = ifelse(SampleID == "LOT1A3_dup_new_protocol", "LOT1A3", SampleID)) %>% 
  column_to_rownames("SampleID")
 
# Export to make a 3D picture in PRIMER
write_tsv(q1.bray.dist.df.source %>% 
            rownames_to_column("SampleID"), "03_Processed_data/16S_glass_slides_q1__bray_dist_dispersal_source.tsv")

write_tsv(data.frame("New_Sample_ID" = row.names(q1.bray.dist.df.source)) %>% 
  left_join(metadata.together) %>%
  mutate(SourceOrEcosystem = ifelse(Substrate == "Glass Slide", 
                                    paste0(Substrate, "_", Ecosystem), 
                                    Substrate)) %>% 
  select(New_Sample_ID, SourceOrEcosystem), 
  "03_Processed_data/16S_glass_slides_q1__bray_dist_dispersal_source_metadata.tsv")



# Do the NMDS
q1.source.NMDS <- metaMDS(as.dist(as.matrix(q1.bray.dist.df.source)), 
                             autotransform = FALSE, k = 3, trymax = 200, trace = 0)


# Merge with our metatdata and create a new column to specify what we want to look at
q1.source.NMDS.meta <- data.frame(q1.source.NMDS$points[,1:2]) %>% 
  rownames_to_column("New_Sample_ID") %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Time_Point, Ecosystem, Treatment)) %>% 
  mutate(Source_Sink = ifelse(Treatment == "Open" | Treatment == "Environmental Litter",
                                     paste0(Treatment, "_", Ecosystem),
                                     Treatment)) 

# Create our visualization of dispersal slides + sources
ggplot(data = q1.source.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(Source_Sink)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1")  +
  ggtitle("Bacteria, glass slide communities + dispersal sources")

# Doing the stats to test differences among groups
adonis(q1.bray.dist.df.source ~ Source_Sink, permutations = 10000, 
       data = q1.source.NMDS.meta, na.rm= TRUE)

pairwise.adonis(q1.bray.dist.df.source, q1.source.NMDS.meta$Source_Sink)

pairwise.adonis2(q1.bray.dist.df.source ~ Source_Sink, data = q1.source.NMDS.meta)

# Looking at the SourceTracker analysis
q1.proportions <- as.data.frame(fread("03_Processed_data/16S_glass_slide__dispersal_source_tracker/mixing_proportions.txt")) %>% 
  rename(Environmental_Grass = EnvG,
         Environmental_Shrub = EnvS) %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Ecosystem), 
            by = c("SampleID" = "New_Sample_ID")) %>% 
  pivot_longer(Environmental_Grass:Unknown, names_to = "Source", values_to = "Proportion")

ggplot(data = q1.proportions) + 
  geom_boxplot(color = "black", aes(x = Ecosystem, y = Proportion, fill = Source)) + 
  theme_classic() + 
  scale_fill_brewer(palette = "Set1") +
  ggtitle("BActeria, Source Tracker dispersal sources")
```

```{r NMDS_dispersal_sources_fungi}
# Load data
glass.slides.ITS <- "03_Processed_data/ITS_glass_slides.rda"
load(glass.slides.ITS)

# Subset our Bray-Curtis distance matrix to include only samples we want to plot and test
q1.ITS.bray.dist.df.source <- q1.ITS.bray.dist.df %>% 
  select(!(contains("LD"))) %>% 
  rownames_to_column("SampleID") %>% 
  filter(!(grepl("LD", SampleID))) %>% 
  column_to_rownames("SampleID")

# Do the NMDS
q1.ITS.source.NMDS <- metaMDS(as.dist(as.matrix(q1.ITS.bray.dist.df.source)), 
                             autotransform = FALSE, k = 2, trymax = 200, trace = 0)

# Merge with our metatdata and create a new column to specify what we want to look at
q1.ITS.source.NMDS.meta <- data.frame(q1.ITS.source.NMDS$points[,1:2]) %>% 
  rownames_to_column("New_Sample_ID") %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Time_Point, Ecosystem, Treatment)) %>% 
  mutate(Source_Sink = ifelse(Treatment == "Open" | Treatment == "Environmental Litter",
                                     paste0(Treatment, "_", Ecosystem),
                                     Treatment))

# Create our visualization of dispersal slides + sources
ggplot(data = q1.ITS.source.NMDS.meta) +
  geom_point(aes(x = MDS1, y = MDS2, color = as.factor(Source_Sink)), size = 4) + 
  theme_classic() +
  scale_color_brewer(palette = "Set1") +
  ggtitle("Fungi, glass slide communities + dispersal sources")

# Doing the stats to test differences among groups
adonis(q1.ITS.bray.dist.df.source ~ Source_Sink, permutations = 10000, 
       data = q1.ITS.source.NMDS.meta, na.rm= TRUE)

pairwise.adonis(q1.ITS.bray.dist.df.source, q1.ITS.source.NMDS.meta$Source_Sink)

pairwise.adonis2(q1.ITS.bray.dist.df.source ~ Source_Sink, data = q1.ITS.source.NMDS.meta)


# Looking at the SourceTracker analysis
q1.ITS.proportions <- as.data.frame(fread("03_Processed_data/ITS_glass_slide__dispersal_source_tracker/mixing_proportions.txt")) %>% 
  rename(Environmental_Grass = EnvG,
         Environmental_Shrub = EnvS) %>% 
  left_join(metadata.together %>% select(New_Sample_ID, Ecosystem), 
            by = c("SampleID" = "New_Sample_ID")) %>% 
  pivot_longer(Air:Unknown, names_to = "Source", values_to = "Proportion")

ggplot(data = q1.ITS.proportions) + 
  geom_boxplot(color = "black", aes(x = Ecosystem, y = Proportion, fill = Source)) + 
  theme_classic() + 
  scale_fill_brewer(palette = "Set1") +
  ggtitle("Fungi, Source Tracker dispersal sources")
```